### 2017-04-20 writing/paper/sections/abstract.tex
Final comments from Henning

\begin{abstract} Two mesh patterns are coincident if they are avoided by the
same set of permutations, and are Wilf-equivalent if they have the same number
of avoiders of each length. We provide sufficient conditions for coincidence
among mesh patterns, whilst also avoiding a longer classical pattern. Using
these conditions we completely classify coincidences between families containing
a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
completely Wilf-classify equivalences of mesh patterns of length 2 whilst also
avoiding the classical pattern 231.

>>>>

\begin{abstract}
    Two mesh patterns are coincident if they are avoided by the same set of
    permutations, and are Wilf-equivalent if they have the same number of
    avoiders of each length. We provide sufficient conditions for coincidence of
    mesh patterns, when only permutations also avoiding a longer classical
    pattern are considered. Using these conditions we completely classify
    coincidences between families containing a mesh pattern of length 2 and a
    classical pattern of length 3. Furthermore, we completely Wilf-classify mesh
    patterns of length 2 inside the class of 231-avoiding permutations.

### 2017-04-20 writing/paper/sections/coincidences.tex
Final comments from Henning

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
\emph{containers}, \ie \(\cont{\lambda} = \cont{\gamma}\).

@@

We now attempt to fully classify coincidences in families characterised by avoidance
of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).

@@

This rule can be understood very in graphical form. In the
pattern in \FigureRef{fig:rule1} we can gain shading in two boxes since if
there is a point in any of these boxes we would get an occurrence of the
dominating pattern \(\perm{3,2,1}\).

\begin{figure}[htb]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
    \end{tikzpicture}

@@

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)

    Now we consider a permutation \(\omega \in \av{\pi}\),
    containing an occurrence of \(m_1\). Consider placing a point in the
    region corresponding to the box \((a,b)\), regardless of where in this
    region we place the point by condition~\ref{prop:dom1:cont} of the
    proposition we create an occurrence of \(\pi\). Therefore there can be no
    points in this region, which implies this is an occurrence of \(m_2\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item   \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq p^{\boks{i,j}_d}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
    implies and occurrence of \(m_2\).
    We consider taking the first branch of every choice.
    Now consider a permutation \(\omega \in \av{\pi}\). Suppose
    \(\omega\) contains \(m_1\) and consider the region corresponding to
    \((a,b)\) in \(R_1\).If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
    of \(m_2\).
    If the region is non-empty, then by \LemmaRef{lem:incdec} and
    condition~\ref{prop:dom2:condc} of the proposition it must contain
    a decreasing subsequence.
    We can choose the topmost point in the region to replace the corresponding
    point in the mesh pattern and the points from the subsequence are now in
    the box southeast of the point. The other conditions allow this to be done
    without points being present in regions that were shaded. Hence there are no
    points in the region corresponding to the box \((a,b)\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

By applying both the First Dominating Pattern Rule and the Second Dominating
Pattern Rule fully explains coincidences of classes of the form
\(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
mesh patterns, confirming experimental observations.

@@

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.

    Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a pattern in \((\sigma,R_1)^\boksall{a,b}\) such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern that contains an
    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

By application of this rule we can reduce the number of classes in
\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
56 classes observed through experimentation.

@@

Intuitively it is easy to see why our previous rules have no
power here. There is nowhere that it is possible to add a single point to
gain an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
a position where addition of an increase, or decrease, provides extra
shading power.

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If the region is empty then we have an occurrence of \(m_2\).
If there is only one point in the region we can choose this to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
Now if we take the other two points as the points in our
mesh pattern then we get an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.
It is also possible to calculate this coincidence by an extension of the Third
Dominating rule, where we allow a sequence of point addition operations, this
is discussed further in the future work section.

>>>>

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie, \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
\emph{containers}, \ie, \(\cont{\lambda} = \cont{\gamma}\).

@@

We now attempt to fully classify coincidences in families characterised by avoidance
of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
finding and explaining all coincidences where \(\av{\{\pi,m\}} = \av{\{\pi,m^\prime\}}\),
where \(\pi\) is a classical pattern of length \(3\).

@@

This rule can be understood in graphical form. In the pattern in
\FigureRef{fig:rule1} we can gain shading in the boxes \((0,2)\), \((2,0)\)
since if there is a point in either of these boxes there would be an occurrence of
the dominating pattern \(\perm{3,2,1}\).

\begin{figure}[htb]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
    \end{tikzpicture}

@@

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now we consider a permutation \(\omega \in \av{\pi}\),
    containing an occurrence of \(m_1\). If there is a point in the
    region corresponding to the box \((a,b)\), then that point, along with the
    points of the occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\).
    Then condition~\eqref{prop:dom1:cont} of the
    proposition implies an occurrence of \(\pi\). Therefore there can be no
    points in this region, which implies the occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom2:cond} Either one of the following hold \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq p^{\boks{i,j}_d}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
    implies an occurrence of \(m_2\).
    We consider taking the first branch of every choice in condition~\eqref{prop:dom2:cond}.
    Now consider a permutation \(\omega \in \av{\pi}\). Suppose
    \(\omega\) contains \(m_1\) and consider the region corresponding to
    \((a,b)\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
    of \(m_2\).
    If the region is non-empty, then by \LemmaRef{lem:incdec} and
    condition~\eqref{prop:dom2:condc} of the proposition it must contain
    a decreasing subsequence.
    We can choose the topmost point in the region to replace the corresponding
    point in the mesh pattern and the points from the subsequence are now in
    the box southeast of the point. The other conditions allow this to be done
    without points being present in regions that were shaded. Hence there are no
    points in the region corresponding to the box \((a,b)\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

Together the First Dominating Pattern Rule and the Second Dominating
Pattern Rule fully explain coincidences of classes of the form
\(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
mesh patterns, confirming experimental observations.

@@

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
    Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a pattern in \((\sigma,R_1)^\boksall{a,b}\) such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern that contains an
    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

This rule reduces the number of classes in
\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
56 classes observed through experimentation.

@@

Intuitively it is easy to see why our previous rules have no
power here. It is impossible to add a single point to a mesh pattern \((12, R)\) and
create an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
a position where addition of an increase, or decrease, provides extra
shading power.

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are coincident in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
patterns that are also coincident to each other by the same reasoning.)

Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If the region is empty then we have an occurrence of \(m_2\).
If there is only one point in the region we can choose this to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
The other two points are an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.
It is also possible to calculate this coincidence by an extension of the Third
Dominating rule, where we allow a sequence of point addition operations, this
is discussed further in the future work section.

### 2017-04-20 writing/paper/sections/conclfut.tex
Final comments from Henning

It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
\textcite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

>>>>

It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
\textcite[Sec.~11]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-04-20 writing/paper/sections/introduction.tex
Do most of Henning's changes
- Change makefile to output to pdf
- Never need compositions of boksall, so removed.
- fixed consistency with av brackets (I hope)
- fixed Henning  using dollar signs in his edits
- changed Henning's long-form cite to nice automatic textcite
- removed duplicate line about considering as set partitions
Final comments from Henning

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
look at classes of pattern where entries can be required to be consecutive in
value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
patterns were studied by \textcite{MR2264071} to establish necessary conditions
for a Schubert variety to be Gorenstein. These definitions are subsumed under
the definition of \emph{mesh patterns}, introduced by
\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
certain permutation statistics.

When considering permutation patterns some of the main questions posed relate to
how and when a pattern is avoided by, or contained in, a arbitrary set of
permutations. Two patterns \(\pi\) and \(\sigma\) are \emph{Wilf-equivalent} if
the number of permutations that avoid \(\pi\) of length \(n\) is equal to the
number of permutations that avoid \(\sigma\) of length \(n\). A stronger
equivalence condition is that of \emph{coincidence}, where the set of
permutations avoiding \(\pi\) is exactly equal to the set of permutations
avoiding \(\sigma\). Avoiding pairs of patterns of the same length with certain
properties has been studied, \textcite{MR2178749} considered avoiding
a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study
avoiding a vincular and a covincular pattern simultaneously in order to achieve
some interesting counting results. However, little work has been done on
avoiding a mesh pattern and a classical pattern simultaneously.

In this work we aim to establish some ground in this field by computing
coincidences and Wilf-classes and calculating some of the enumerations of
avoiders of a mesh pattern of length 2 and a classical pattern of length 3. We
begin by establishing coincidences between mesh patterns of length 2 while
avoiding a classical pattern of length 3; this is used to establish sufficient
conditions for coincidence. We then establish Wilf-equivalence classes of these
coincidence classes where the classical pattern 231.

>>>>

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
look at classes of patterns where entries can also be required to be consecutive in
value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
patterns were studied by \textcite{MR2264071} to establish necessary conditions
for a Schubert variety to be Gorenstein. These definitions are subsumed under
the definition of \emph{mesh patterns}, introduced by
\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
certain permutation statistics.

When considering permutation patterns some of the main questions posed relate to
how and when a pattern is avoided by, or contained in, an arbitrary set of
permutations. Two patterns \(\pi\) and \(\sigma\) are \emph{Wilf-equivalent} if
the number of permutations that avoid \(\pi\) of length \(n\) is equal to the
number of permutations that avoid \(\sigma\) of length \(n\). A stronger
equivalence condition is that of \emph{coincidence}, where the set of
permutations avoiding \(\pi\) is exactly equal to the set of permutations
avoiding \(\sigma\). Avoiding pairs of patterns of the same length with certain
properties has been studied, \textcite{MR2178749} considered avoiding
a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study
avoiding a vincular and a covincular pattern simultaneously in order to achieve
some several counting results. However, little work has been done on
avoiding a mesh pattern and a classical pattern simultaneously.

In this work we aim to establish some ground in this field by computing
coincidences and Wilf-classes and calculating some of the enumerations of
avoiders of a mesh pattern of length 2 and a classical pattern of length 3. We
begin by establishing coincidences between mesh patterns of length 2 while
avoiding a classical pattern of length 3; this is used to establish sufficient
conditions for coincidence. We then establish Wilf-equivalence classes of these
coincidence classes where the classical pattern is \perm{2,3,1}.

### 2017-04-20 writing/paper/sections/wilf.tex
Results of Discussions with Henning
- Changed example to be about subpatterns, may need more words
- Removed second description of one example
Final comments from Henning

Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
must only use these symmetries when they preserve the dominating pattern, if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

@@

\begin{equation*}
    \scriptvar = \varepsilon \sqcup
\decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
If the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left
section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}\\
    \var(x) &= \frac{C(x)}{1+xC(x)}
\end{align*} % TODO Do we need both representations?

@@

\begin{note}
  The avoiders of the pattern \(q = \mperm{\perm{2,3,1}}{(1,0),(1,1),(1,2),(1,3)},
\textpattern{}{2,3,1}{1/0,1/1,1/2,1/3}\) in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions
of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01})
\end{note}
\begin{example}
    The permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
\end{example}

@@

This particular example is interesting as it shows that both methods can be used in tandem to establish a coincidence class.

@@

\noindent
Here \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\).
Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\). It can
once again be decomposed around the leftmost point
\begin{equation*}
    \scriptvar^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_1\setminus\varepsilon}{}{}{\scriptvar^\prime_1}
\end{equation*}
This is a complete decomposition of avoiders of \(m_1\). Now we
look at an avoider of \(m_2\), this time decomposition is around the leftmost point
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptvar_2}
\end{equation*}

>>>>

Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
must only use symmetries that preserve the dominating pattern(s), if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

@@

\begin{equation*}
    \scriptvar = \varepsilon \sqcup
\decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
If the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left
section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1-x(C(x)-1)}
\end{align*}

@@

\begin{note}
  The avoiders of the mesh pattern \(q = %= \mperm{\perm{2,3,1}}{(1,0),(1,1),(1,2),(1,3)}
\textpattern{}{2,3,1}{1/0,1/1,1/2,1/3}\) in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions
of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01}).
\end{note}
\begin{example}
    The permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
\end{example}

@@

% This particular example is interesting as it shows that both methods can be used in tandem to establish a coincidence class.

@@

\noindent
Here \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\).
Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\). It can
once again be decomposed around the leftmost point
\begin{equation*}
    \scriptvar^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_1\setminus\varepsilon}{}{}{\scriptvar^\prime_1}
\end{equation*}
This is a complete decomposition of avoiders of \(m_1\). Now we
look at an avoider of \(m_2\), decomposed around the leftmost point
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptvar_2}
\end{equation*}
